### 2022-03-22 paper/output/document/thesis_5_discussion.tex
finished another read thrgh

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In the above chapters, I presented an implementation of automated Kantian ethics that, given an appropriately
represented maxim and sufficient factual background, can classify a maxim as obligatory, permissible, 
or prohibited. In this chapter, I discuss the philosophical limitations and implications of this 
work.

@@

In Section \ref{amapossible}, I discuss theoretical objections to automated ethics, derived from the
common philosophical intuition that there is no algorithm for ethics. I explore concrete versions of 
this objection and make a theoretical argument for the possibility of the kind of automated ethics 
I implement in this thesis. Finally, in Section \ref{relatedwork}, I situate my project among related work.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that, given
sufficient factual or ``common sense'' background, can take 
as input the logical representation of a maxim and return its moral status (if it is obligatory, 
prohibited, or permissible). My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. \begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent. This ethics engine passes a moral dilemma 
through an input parser, applies the automated categorical imperative test, and finally processes the 
output using an output parser, producing a prescription for action. I contribute the automated categorical 
imperative component.} \label{fig:AIengine}
\end{figure} Figure \ref{fig:AIengine} depicts the workflow of this example ethics engine.

In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work must develop such input and output parsers.

One of the biggest challenges for an ethics engine is the development of an input parser. An input parser 
for my implementation of automated Kantian ethics must translate a complex real-world situation into a flat, logical representation.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As mentioned in Section \ref{whatisamaxim},
there is robust debate on the circumstances that should be considered when formulating a maxim, 
inspired by a common criticism of Kantian ethics called the tailoring objection. Recall that the 
tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

@@

While the above account explains how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. However an action is turned into a maxim for my system 
to process, whether manually as I did in Chapter \ref{applications} or using an automatic input 
parser, this transformation must be a good-faith attempt to capture the principle of action. 
Correctly formulating a maxim requires what Kantians call ``practical judgement,'' or
common sense reasoning and factual background \citep{oneilluniversallaws}. Returning to the example above, the fact that being 
unemployed may contribute to one's desire to steal another's job is a consequence of practical
judgement, not just pure reason alone. Translating everyday situations into appropriate maxims is
the bulk of the reasoning that a Kantian human being does when making decisions. Automating this
reasoning requires endowing a machine with common sense.

The challenge of formulating a maxim is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser by supervising the operation of an AI agent. When the agent stumbles onto
an ethical dilemma, the human could take over, formulate
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. Alternatively, AI developers can identify expected ethical dilemmas 
that the machine may face in advance and hardcode their own judgements for these dilemmas. 
For proponents of the ``human-in-the-loop'' model of AI ethics, in which
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
Both of these solutions imply that the outcome of the universalizability test will depend on how the 
human formulates the maxim; if the human puts garbage into the test, the test will return garbage out.

@@

These interpretations could inspire heuristics to classify circumstances as morally 
relevant. For example, one such attempt could define a moral closeness relation between an action, a 
goal, and circumstances. This heuristic could define morally relevant circumstances as those that 
reach a certain closeness threshhold with the action and the goal. Another possible heuristic could 
define some set of morally important entities, and classify morally relevant circumstances as those
that involve morally important entities. I propose a potential machine-learning based approach which formulates
maxims based on a training set of appropriately formulated maxims in Section \ref{amapossible}. This 
approach mimics how human beings formulate maxims; we use common sense and prior situational, 
factual, and ethical knowledge to isolate our principle of action. Determining morally relevant circumstances, 
either using heuristics or human involvement, is a ripe area for future work.

Once the input has been parsed, either by a human or a machine, into a sentence in my logic, my 
project can evaluate its moral status using my implementation of 
the FUL. Concretely, my project returns a value indicating if the maxim is obligatory, permissible, 
or prohibited. The maxim is prohibited if it fails the universalizability test, permissible if it passes, and obligatory 
if its negation fails the universalizability test. All three of these properties require testing if a 
certain theorem holds or not in my logic, a calculation that I demonstrate in Section \ref{testing}. 
Testing these properties requires that my system have a database of common sense or factual background. 
Different applications of my system may require different factual background (e.g. a self-driving car 
needs to know traffic regulations), so this common sense database will need to be application 
specific. As demonstrated in the examples in Chapter \ref{applications}, my system can produce sophisticated 
judgements with relatively little situational context. Thus, while my automated categorical imperative
requires some common sense and factual background, Chapter \ref{applications} demonstrates that automating
this common sense is less daunting than it seems.

@@

I do not argue that computational ethics uncovers philosophical insights that humans are always incapable 
of reaching. In some cases, computational ethics may automate calculations too long and tedious for 
any human philosopher to complete. Insights like the one about well-formed maxims, however, could 
have been discovered by a human being but are much easier to reach with computational tools. Computational tools prompt philosophers 
to ask new questions that lead to insights, and can thus serve as another tool in a philosopher's arsenal, like a 
thought experiment or counterexample.

@@

Computational ethics is at its infancy. The use of theorem provers in mathematics is only now beginning 
to make headway \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers to automate deontic logic occurred in the last
few years. The 
fact that this nascent technology is already helping humans reach non-trivial ethical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our expectations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it could
become as commonplace in philosophy departments as reflective equilibrium. Just as computational tools
have amplified progress in healthcare and drug discovery, computational ethics has the potential to enable
great philosophical progress.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automating Everyday Practical Reason \label{ordinarypeople}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Sections \ref{AIethics} and \ref{computationalethics}, I outline how automated ethics can guide
artificial agents and philosophers respectively. This raises a natural question: can automated
ethics guide ordinary human beings, not just academic philosophers, as we navigate the world and face ethical
dilemmas? Some may hope (or worry) that automated ethics could render ethical reasoning obsolete. In 
this section, I argue that while computers should not replace human ethical reasoning entirely,
they can supplement and improve our ethical reasoning. I argue for a kind of human-computer symbiosis
in which computers offer ethical advice, arguments for particular moral judgements, and speed up moral
calculations without subverting human ethical reasoning entirely \citep{licklider}.

@@

The above model of human-computer symbiosis demonstrates how computational ethics can augment human
ethical reasoning without replacing it. When deliberating over moral dilemmas, ordinary people can 
turn to computational tools for advice, like an ``Ask an Ethicist'' column. If we appeal to philosophically
faithful computational ethics like that implemented in this thesis, then this advice will synthesize 
decades of philosophical progress and is thus a way to apply philosophers' insights to ordinary life. 
Moreover, just as they do for philosophers, computers can help ordinary people approach ethical questions from a 
different perspective. Even interacting with my system requires the user to consider the action's maxim, 
which includes the circumstances, act, and goal. Making these components of action precise already changes
the user's perspective. Just as computational ethics can serve as a tool for academic philosophers to 
ask new questions and achieve greater precision, it can do the same for ordinary human beings navigating
the world. Moreover, it also offers another way for the general public to access professional philosophy
insights, and thus carries potential to improve our everyday reasoning.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Theoretical Objections to Automating Kantian Ethics \label{amapossible}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Many philosophers cringe at the idea that a computer could perform ethical reasoning or that the 
categorical imperative could provide an algorithm for moral judgement. For example, Rawls asserts, 
``it is a serious misconception to think of the CI-procedure as an algorithm intended to yield, 
more or less mechanically, a correct judgment. There is no such algorithm, and Kant knows this'' \citep[166]{rawlslectures}. 
Ebels-Duggan also claims, ``no one supposes that the Categorical Imperative provides a mechanical 
algorithm that delivers all by itself a complete account of what we ought to do in any given 
situation'' \citep[174]{ebelsduggan}. However unmechanical ethical reasoning
may seem, these claims are not obvious and require further justification. Philosophers who believe 
that mental activity completely determines moral reasoning must explain why computers can, in theory, 
simulate certain mental processes like arithmetic and language, but cannot perform ethical reasoning. 
Without a soul or God-based account of ethical reasoning, it is not obvious that it is theoretically 
impossible to automate ethical reasoning. After all, computers may eventually learn to simulate human 
mental activity entirely, as shown by progress in brain simulation \citep{brainsimulation}. Skeptics
about automated ethics must explain why ethics is any different from other automated human activity.

@@

Another challenge for automated Kantian ethics is that the FUL test requires that
a maxim be given as input. O'Neill notes that the test assumes ``that agents will have certain tentative 
plans, proposals and policies which they can consider, revise or reject or endorse and pursue'' \citep[343]{oneilluniversallaws}.
The FUL evaluates the moral worth of a maxim given as input, where this potential maxim is generated by 
the choices that an agent is faced with. As I argue in Section \ref{AIethics}, determining this potential maxim is a challenge for both human
and automated reasoners. Kant even claims that the difficulty of determining an agent's potential maxim, which is their
own, subjective understanding of their principle of action, is a reason that we may never be able 
to know if the morally worthy action has been performed \cite[345]{oneilluniversallaws}. Reasoners are faced with
choices between potential actions and must determine the maxim, or principle, underlying each potential action.
This is equivalent to a ``mapping'' problem: agents are given situations or dilemmas as input and must map
these to maxims.

@@

instead of a derivation tool also explains the response to the tailor objection presented in Section \ref{AIethics}, that the FUL cannot 

@@

Under Tafani's understanding of the categorical imperative, not only is automated moral reasoning

@@

by double checking the moral intuitions that we take for granted. A computer with no common sense or prior knowledge

@@

In this paper, I use a top-down approach to formalize Kantian ethics. There is a long line of work 
automating other ethical theories, like consequentialism \citep{util1, util2} or particularism 
\citep{particularism1, particularism2}. I choose to implement Kantian ethics because, as argued in 
Section \ref{whykant}, it is the most formal and least data-intensive of the three major ethical 
traditions. Kantian ethics is a deontological, or rule based ethic, and there is prior work 
implementing other deontological theories \citep{dde, deon1, deon2}.

@@

The implementation of this paper was inspired by and builds on Benzmüller, Parent, and Farjami's 
foundational work with the LogiKEy framework for machine ethics, which includes their implementation 
of DDL in Isabelle \citep{BFP, logikey}. The LogiKEy project has been used to study metaphysics 
\citep{godel, metaphysics1}, law \citep{constitution}, and ethics \citep{gewirth}, but not 
Kant's categorical imperative.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Conclusion%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this thesis, I present a proof-of-concept implementation of automated Kantian ethics. My system
takes as input a potential action, appropriately represented, and can prove that it is obligated, 
prohibited or permissible. I represent Kant's Formula of Universal Law in a deontic logic and 
implement this logic in the Isabelle/HOL interactive theorem prover, which can automatically prove or 
refute theorems in my custom logic. I also contribute a testing framework that demonstrates that
my implementation of Kantian ethics is more faithful to philosophical literature than two other 
potential implementations. My completed system can, when given appropriate factual background, make
philosophically mature judgements about complex moral dilemmas. This work is one step towards building 
morally sophisticated artifical agents.

@@

%:%60=25%:%
%:%72=27%:%
%:%73=28%:%
%:%74=29%:%
%:%75=30%:%
%:%76=31%:%
%:%77=32%:%
%:%78=33%:%
%:%79=34%:%
%:%80=35%:%
%:%81=36%:%
%:%82=37%:%
%:%83=38%:%
%:%84=39%:%
%:%85=40%:%
%:%86=41%:%
%:%87=42%:%
%:%88=43%:%
%:%89=44%:%
%:%90=45%:%
%:%91=46%:%
%:%92=47%:%
%:%93=48%:%
%:%94=49%:%
%:%95=50%:%
%:%96=51%:%
%:%97=52%:%
%:%98=53%:%
%:%99=54%:%
%:%100=55%:%
%:%101=56%:%
%:%102=57%:%
%:%103=58%:%
%:%104=59%:%
%:%105=60%:%
%:%106=61%:%
%:%107=62%:%
%:%108=63%:%
%:%109=64%:%
%:%110=65%:%
%:%111=66%:%
%:%112=67%:%
%:%113=68%:%
%:%114=69%:%
%:%115=70%:%
%:%116=71%:%
%:%117=72%:%
%:%118=73%:%
%:%119=74%:%
%:%120=75%:%
%:%121=76%:%
%:%122=77%:%
%:%123=78%:%
%:%124=79%:%
%:%125=80%:%
%:%126=81%:%
%:%127=82%:%
%:%128=83%:%
%:%129=84%:%
%:%130=85%:%
%:%131=86%:%
%:%132=87%:%
%:%133=88%:%
%:%134=89%:%
%:%135=90%:%
%:%136=91%:%
%:%137=92%:%
%:%138=93%:%
%:%139=94%:%
%:%140=95%:%
%:%141=96%:%
%:%142=97%:%
%:%143=98%:%
%:%144=99%:%
%:%145=100%:%
%:%146=101%:%
%:%147=102%:%
%:%148=103%:%
%:%149=104%:%
%:%150=105%:%
%:%151=106%:%
%:%152=107%:%
%:%153=108%:%
%:%154=109%:%
%:%155=110%:%
%:%156=111%:%
%:%157=112%:%
%:%158=113%:%
%:%159=114%:%
%:%160=115%:%
%:%161=116%:%
%:%162=117%:%
%:%163=118%:%
%:%164=119%:%
%:%165=120%:%
%:%166=121%:%
%:%167=122%:%
%:%168=123%:%
%:%169=124%:%
%:%170=125%:%
%:%171=126%:%
%:%172=127%:%
%:%173=128%:%
%:%174=129%:%
%:%175=130%:%
%:%176=131%:%
%:%177=132%:%
%:%178=133%:%
%:%179=134%:%
%:%180=135%:%
%:%181=136%:%
%:%182=137%:%
%:%183=138%:%
%:%184=139%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In the above chapters, I presented an implementation of automated Kantian ethics that, given an appropriately
represented maxim and sufficient factual background, can classify a maxim as obligatory, permissible, 
or prohibited. In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this chapter, I discuss the philosophical limitations and implications of this 
work.

@@

In Section \ref{amapossible}, I discuss theoretical objections to automated ethics, derived from the
common philosophical intuition that there is no algorithm for ethics. I explore concrete versions of 
this objection and make a theoretical argument for the possibility of the kind of automated ethics 
I implement in this thesis. Finally, in Section \ref{relatedwork}, I situate my project among related work.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that can
evaluate the moral status of a maxim. My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, when an AI agent faces a moral dilemma in some internal representation, it could pass it 
to an input parser that translates the dilemma into a maxim in my logic. An output parser could then 
translate my system's output into a prescription for action that the AI agent could act on. \begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent. This ethics engine passes a moral dilemma 
through an input parser, applies the automated categorical imperative test, and finally processes the 
output using an output parser, producing a prescription for action. I contribute the automated categorical 
imperative component.} \label{fig:AIengine}
\end{figure} Figure \ref{fig:AIengine} depicts the workflow of this example ethics engine. In order 
for my system to guide an AI agent using this workflow, future work must develop such input and output parsers.

One of the biggest challenges for an ethics engine is the development of an input parser. An input parser 
for my implementation of automated Kantian ethics must translate a complex real-world situation into a flat 
circumstances, act, goal triple.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As introduced in Section \ref{whatisamaxim},
there is robust debate on the circumstances that should be considered when formulating a maxim, 
inspired by a common criticism of Kantian ethics called the tailoring objection. Recall that the 
tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

@@

While the above account explains how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. However an action is turned into a maxim for my system 
to process, whether manually as I did in Chapter \ref{applications} or using an automatic input 
parser, this transformation must be a good-faith attempt to capture the principle of action. 
Such a good-faith attempt to formulate a maxim requires what Kantians call ``practical judgement,'' or
common sense reasoning and factual background \citep{oneilluniversallaws}. Returning to the example above, the fact that being 
unemployed may contribute to one's desire to steal another's job is a consequence of practical
judgement, not just pure reason alone. Automating the formulation of a maxim requires
endowing a machine with common sense.

Translating everyday situations into appropriate maxims is the bulk of the reasoning that a Kantian 
human being does when making decisions, so it is unsurprising that formulating a maxim 
is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser by supervising the operation of an AI agent. When the agent stumbles onto
an ethical dilemma, the human could take over, formulate
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. Alternatively, AI developers can identify expected ethical dilemmas 
that the machine may face in advance and hardcode their own judgements for these dilemmas. 
For proponents of the ``human-in-the-loop'' model of AI ethics, in which
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
Both of these solutions imply that the outcome of the universalizability test will depend on how the 
human formulates the maxim; if the human puts garbage into the test, the test will return garbage out.

@@

These interpretations could inspire heuristics to classify circumstances as morally 
relevant. For example, one such attempt could define a moral closeness relation between an act, a 
goal, and circumstances. This heuristic could define morally relevant circumstances as those that 
reach a certain closeness threshhold with the act and the goal. Another possible heuristic could 
define some set of morally important entities, and classify morally relevant circumstances as those
that involve morally important entities. I propose a potential machine-learning based approach which formulates
maxims based on a training set of appropriately formulated maxims in Section \ref{amapossible}. This 
approach mimics how human beings formulate maxims; we use common sense and prior situational, 
factual, and ethical knowledge to isolate our principle of action. Determining morally relevant circumstances, 
either using heuristics or human involvement, is a ripe area for future work.

Once the input has been parsed, either by a human or a machine, into a sentence in my logic, my 
project can evaluate its moral status using my implementation of 
the FUL. Concretely, my project returns a value indicating if the maxim is obligatory, permissible, 
or prohibited. The maxim is prohibited if it fails the universalizability test, permissible if it passes, and obligatory 
if its negation fails the universalizability test. All three of these properties require testing if a 
certain theorem holds or not in my logic, a calculation that I demonstrate in Section \ref{testing}. 
Testing these properties requires that my system have a database of common sense or factual background. 
Different applications of my system may require different factual background (e.g. a self-driving car 
needs to know traffic regulations), so this common sense database will need to be application 
specific. As demonstrated in the examples in Chapter \ref{applications}, my system can produce sophisticated 
judgements with relatively little situational context. Thus, while my automated categorical imperative system
requires some common sense and factual background, Chapter \ref{applications} demonstrates that automating
this common sense is less daunting than it seems.

@@

In some cases, computational ethics may automate calculations too long and tedious for 
any human philosopher to complete. Insights like the one about well-formed maxims, however, could 
have been discovered by a human being but are much easier to reach with computational tools. I do not 
argue that computational ethics uncovers philosophical insights that humans are always incapable 
of reaching. Computational tools prompt philosophers 
to ask new questions that lead to insights, and can thus serve as another tool in a philosopher's arsenal, like a 
thought experiment or counterexample.

@@

Computational ethics is at its infancy. Theorem provers are only now beginning to make headway in 
mathematics \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers to automate deontic logic occurred in the last
few years. The fact that this nascent technology is already helping humans reach non-trivial ethical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our expectations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it could
become as commonplace in philosophy departments as reflective equilibrium. Just as computational tools
have amplified progress in healthcare and drug discovery, computational ethics has the potential to enable
great philosophical progress.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automating Everyday Practical Reason \label{ordinarypeople}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Sections \ref{AIethics} and \ref{computationalethics}, I outline how automated ethics can guide
artificial agents and philosophers respectively. This raises a natural question: can automated
ethics guide all human beings, not just academic philosophers, as we navigate the world and face ethical
dilemmas? Some may hope (or worry) that automated ethics could render everyday ethical reasoning obsolete. In 
this section, I argue that while computers should not replace human ethical reasoning entirely,
they can supplement and improve our ethical reasoning. I argue for a kind of human-computer symbiosis
in which computers offer ethical advice, arguments for particular moral judgements, and speed up moral
calculations without subverting human ethical reasoning entirely \citep{licklider}.

@@

The above model of human-computer symbiosis demonstrates how computational ethics can augment human
ethical reasoning without replacing it. When deliberating over moral dilemmas, ordinary people can 
turn to computational tools for advice, like an ``Ask an Ethicist'' column. If we appeal to philosophically
faithful computational ethics like that implemented in this thesis, then this advice will synthesize 
decades of philosophical progress and is thus a way to apply philosophers' insights to ordinary life. 
Moreover, just as they do for philosophers, computers can help ordinary people approach ethical questions from a 
different perspective. Even interacting with my system requires the user to consider the action's maxim, 
which includes the circumstances, act, and goal. Making these components of action precise already changes
the user's perspective. Just as computational ethics can serve as a tool for academic philosophers to 
ask new questions and achieve greater precision, it can do the same for all human beings navigating
the world. Moreover, it also offers another way for the general public to access professional philosophy's
insights, and thus carries potential to improve our everyday reasoning.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Theoretical Objections to Automating Kantian Ethics \label{amapossible}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Many philosophers cringe at the idea that a computer could perform ethical reasoning or that the 
categorical imperative could provide an algorithm for moral judgement. For example, Rawls asserts, 
``it is a serious misconception to think of the CI-procedure as an algorithm intended to yield, 
more or less mechanically, a correct judgment. There is no such algorithm, and Kant knows this'' \citep[166]{rawlslectures}. 
Ebels-Duggan also claims, ``no one supposes that the Categorical Imperative provides a mechanical 
algorithm that delivers all by itself a complete account of what we ought to do in any given 
situation'' \citep[174]{ebelsduggan}. However unmechanical ethical reasoning
may seem, these claims are not obvious and require further justification. Philosophers who believe 
that mental activity completely determines moral reasoning must explain why computers can, in theory, 
simulate certain mental processes like arithmetic and language, but cannot perform ethical reasoning. 
Without a soul or God-based account of ethical reasoning, it is not obvious that it is theoretically 
impossible to automate ethical reasoning. After all, computers may eventually learn to simulate human 
mental activity entirely, as shown by progress in brain simulation \citep{brainsimulation}. Skeptics
about automated ethics must explain why ethics is any different from other automated human activity.

@@

Another challenge for automated Kantian ethics is that the FUL test requires that
a maxim be given as input. O'Neill notes that the test assumes ``that agents will have certain tentative 
plans, proposals and policies which they can consider, revise or reject or endorse and pursue'' \citep[343]{oneilluniversallaws}.
The FUL evaluates the moral worth of a maxim given as input, where this potential maxim is generated by 
the choices that an agent is faced with. As I argue in Section \ref{AIethics}, determining this potential maxim is a challenge for both human
and automated reasoners. Kant claims that determining an agent's potential maxim is so difficult because a maxim is their
prsonal understanding of their principle of action, and so we may never be able 
to know if the morally worthy action has been performed \cite[345]{oneilluniversallaws}. Reasoners are faced with
choices between potential actions and must determine the maxim, or principle, underlying each potential action.
This is equivalent to a ``mapping'' problem: agents are given situations or dilemmas as input and must map
these to maxims.

@@

instead of a derivation tool also explains the response to the tailoring objection presented in Section \ref{AIethics}, that the FUL cannot 

@@

Under Tafani's and O'Neill's understandings of the categorical imperative, not only is automated moral reasoning

@@

by double-checking the moral intuitions that we take for granted. A computer with no common sense or prior knowledge

@@

In this paper, I use a top-down approach to formalize Kantian ethics. There is a long line of work 
automating other ethical theories, like consequentialism \citep{util2, util1} or particularism 
\citep{particularism1, particularism2}. I choose to implement Kantian ethics because, as argued in 
Section \ref{whykant}, it is the most formal and least data-intensive of the three major ethical 
traditions. Kantian ethics is a deontological, or rule based ethic, and there is prior work 
implementing other deontological theories \citep{deon1, deon2, dde}.

@@

The implementation of this paper was inspired by and builds on Benzmüller, Parent, and Farjami's 
foundational work with the LogiKEy framework for machine ethics, which includes their implementation 
of DDL in Isabelle \citep{logikey, BFP}. The LogiKEy project has been used to study metaphysics 
\citep{godel, metaphysics1}, law \citep{constitution}, and ethics \citep{gewirth}, but not 
Kant's categorical imperative.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Conclusion%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this thesis, I present a proof-of-concept implementation of automated Kantian ethics. My system
takes as input a potential action, appropriately represented, and can prove that it is obligated, 
prohibited or permissible. I represent Kant's Formula of Universal Law in a deontic logic and 
implement this logic in the Isabelle/HOL interactive theorem prover, which can automatically prove or 
refute theorems in my custom logic. I also contribute a testing framework that demonstrates that
my implementation of Kantian ethics is more faithful to philosophical literature than two other 
potential implementations. My completed system can, when given appropriate factual background, make
philosophically mature judgements about complex moral dilemmas. This work is one step towards building 
morally sophisticated artifical agents.

@@

%:%52=24%:%
%:%61=26%:%
%:%73=27%:%
%:%74=28%:%
%:%75=29%:%
%:%76=30%:%
%:%77=31%:%
%:%78=32%:%
%:%79=33%:%
%:%80=34%:%
%:%81=35%:%
%:%82=36%:%
%:%83=37%:%
%:%84=38%:%
%:%85=39%:%
%:%86=40%:%
%:%87=41%:%
%:%88=42%:%
%:%89=43%:%
%:%90=44%:%
%:%91=45%:%
%:%92=46%:%
%:%93=47%:%
%:%94=48%:%
%:%95=49%:%
%:%96=50%:%
%:%97=51%:%
%:%98=52%:%
%:%99=53%:%
%:%100=54%:%
%:%101=55%:%
%:%102=56%:%
%:%103=57%:%
%:%104=58%:%
%:%105=59%:%
%:%106=60%:%
%:%107=61%:%
%:%108=62%:%
%:%109=63%:%
%:%110=64%:%
%:%111=65%:%
%:%112=66%:%
%:%113=67%:%
%:%114=68%:%
%:%115=69%:%
%:%116=70%:%
%:%117=71%:%
%:%118=72%:%
%:%119=73%:%
%:%120=74%:%
%:%121=75%:%
%:%122=76%:%
%:%123=77%:%
%:%124=78%:%
%:%125=79%:%
%:%126=80%:%
%:%127=81%:%
%:%128=82%:%
%:%129=83%:%
%:%130=84%:%
%:%131=85%:%
%:%132=86%:%
%:%133=87%:%
%:%134=88%:%
%:%135=89%:%
%:%136=90%:%
%:%137=91%:%
%:%138=92%:%
%:%139=93%:%
%:%140=94%:%
%:%141=95%:%
%:%142=96%:%
%:%143=97%:%
%:%144=98%:%
%:%145=99%:%
%:%146=100%:%
%:%147=101%:%
%:%148=102%:%
%:%149=103%:%
%:%150=104%:%
%:%151=105%:%
%:%152=106%:%
%:%153=107%:%
%:%154=108%:%
%:%155=109%:%
%:%156=110%:%
%:%157=111%:%
%:%158=112%:%
%:%159=113%:%
%:%160=114%:%
%:%161=115%:%
%:%162=116%:%
%:%163=117%:%
%:%164=118%:%
%:%165=119%:%
%:%166=120%:%
%:%167=121%:%
%:%168=122%:%
%:%169=123%:%
%:%170=124%:%
%:%171=125%:%
%:%172=126%:%
%:%173=127%:%
%:%174=128%:%
%:%175=129%:%
%:%176=130%:%
%:%177=131%:%
%:%178=132%:%
%:%179=133%:%
%:%180=134%:%
%:%181=135%:%
%:%182=136%:%
%:%183=137%:%
%:%184=138%:%
